Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-03-02 cluster_paper/ms.tex
update text

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: The photometric data are from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars having not yet reached the
    main-sequence.
    {\it Right}: Solar-metallicity PARSEC isochrones are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
    The main-sequence turn-off location and evolved stars suggests a
    cluster age slightly older than the Pleiades.  The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.  The halo members show a roughly
consistent main-sequence, perhaps with greater scatter.  One
possible explanation for the additional scatter could be that the halo
is more contaminated by field stars.  
For instance, $\approx5$ evolved halo stars are visible in the
field-star portion of the red giant branch (RGB).
Given the expected age of the system, these stars must be field
interlopers.

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening.  The halo is reported to span
20$^\circ$ on-sky, and varies in position from about $b=-12^\circ$ to
$b=-20^\circ$, with the stars closest to the galactic plane also being
further from the Sun by up to 200\,pc (Figure~\ref{fig:gaia6d}).
Some empirical evidence for differential reddening is discussed
further in Appendix~\ref{app:diffred}. Based on the available
evidence, we expect that both field star contamination and
differential reddening could play a role.
A third possibility that we have not explored is whether the binary
fraction could differ in spatially different regions of the cluster.

In the right panel of Figure~\ref{fig:hr}, we compare the observed
photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}} to interpolate these isochrones
at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied the
reddening correction described in Section~\ref{subsec:tess}. This
entailed by including using our corrected value of
$E(B-V)=0.102\pm0.019$, assuming $R_V=3.1$, and applying the
\citet{odonnell_1994} SED-dependent reddening law.
The data and models generally agree for masses above
approximately 0.45\,$M_\odot$.  Below this mass, at colors redder than
Bp-Rp$\approx2.2$, the data and models diverge.  The MIST isochrones show a
comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
both starspots and incomplete molecular line lists have been proposed
\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the 
PARSEC isochrones since they diverge at slightly lower mass than the
MIST models, due to the temperature-opacity calibration
discussed by \citet{chen_improving_2014}.
For purposes of age-dating the cluster, we prefer to simply focus on
the main-sequence turn-off (MSTO), since it is where the models have
maximal predictive power.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} shows
empirically that they seem to be: stars past the turnoff in both the
halo and core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} suggests the precision with which the claim can be
made.  The MSTO stars are consistent with the 178 and 316 Myr models,
but are bluer than the 100 Myr model.  The RGB stars are consistent
only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$) model.  Based on
the assumption that the width of the MSTO can be attributed to binary
stars, we are most interested in the blue edge. This seems most
compatible with the 178 Myr model.

% Our photometric age-dating procedure is as follows.  We selected all
% stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
% excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
% outliers from the rest of the cluster.  Using an isochrone grid
% spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
% spacing, we then numerically evaluated the $\chi^2$ fit for each
% possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
% percentiles are used as summary statstics.
% 
% % NOTE: probably asymmetric.
% The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
% For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
% yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
% the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars not having yet reached the
    main-sequence.
    {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$.
    The arrow represents the average reddening correction applied
    to the models.
    The main-sequence turn-off location and evolved stars suggests a
    cluster age slightly older than the Pleiades (dark blue line).  The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.  The halo members show a roughly
consistent main-sequence, with somewhat greater scatter.  One
possible explanation for the additional scatter could be that the
halo is more contaminated by field stars.  For instance, $\approx5$
evolved halo stars are visible in the field-star portion of the red
giant branch (RGB).  Given the age of the system, these stars must be
field interlopers.

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The halo
is reported to span 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 200\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed further in Appendix~\ref{app:diffred}. Based on
the available evidence, we expect that both field star contamination
and differential reddening could play a role in the larger scatter of
the halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

In the right panel of Figure~\ref{fig:hr}, we compare the observed
photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
the reddening correction described in Section~\ref{subsec:tess}.  We
assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applyied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.  For the metallicity, we considered a
range of super and sub-solar metallicities to fit as much of the locus
from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
super-solar $[M/H]=0.06$.  Sub-solar metallicities led to model
predictions too blue along the main sequence by $\approx$0.1\,mag.
Our adopted metallicity agrees with that found by
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for
the cluster, we prefer super-solar based on the goodness of fit from
the photometry.

Overall, the data and models agree for masses above
approximately 0.45\,$M_\odot$.  However below this mass, the data and
models diverge at colors redder than Bp-Rp$\approx2.2$.  The MIST
isochrones showed a comparable disagreement \citep{choi_mesa_2016}.
Explanations invoking both starspots and incomplete molecular line
lists have been proposed \citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the PARSEC isochrones since they diverge at slightly lower
mass than the MIST models, due to the temperature-opacity calibration
implemented by \citet{chen_improving_2014}.  For purposes of
age-dating the cluster, we prefer to simply focus on the main-sequence
turn-off (MSTO), since it is where the models have maximal predictive
power.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} shows
empirically that they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim can be
made.  The MSTO stars are consistent with the 178 and 316 Myr models,
but are bluer than the 100 Myr model.  The RGB stars are consistent
only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$) model.  Based on
the assumption that the width of the MSTO can be attributed to binary
stars, we are most interested in the blue edge. This appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for a more precise model-dependent
comparison.

% FIXME: do you want to do the procedure below? Seems like overkill
% given the Cummings study.
% Our photometric age-dating procedure is as follows.  We selected all
% stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
% excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
% outliers from the rest of the cluster.  Using an isochrone grid
% spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
% spacing, we then numerically evaluated the $\chi^2$ fit for each
% possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
% percentiles are used as summary statstics.
% 
% % NOTE: probably asymmetric.
% The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
% For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
% yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
% the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.
